T-reX: Quantifying waste and material footprints in current and future Life Cycle Assessment (LCA) databases

### 2024-04-02 latex_files/methodology.tex
update

In addition to the waste categories, T-reX includes numerous material demand categories, which are based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}. The CRM list contains 30 materials that are considered critical to the EU economy and are at risk of supply disruption. Several further materials of interest to the authors were added to the search list, including helium, electricity, petroleum, sand, water, and natural gas. The identity of the materials considered and their categorical groupings are easily customisable by the user. The full list of 59 materials included in the default configuration is provided in the supplementary material in ~\autoref{sec:supplementary}.

@@

\begin{figure}[H]
    \centering
    \caption{Workflow of T-reX. Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
    \includegraphics[width=0.5\linewidth]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
\end{figure}

@@

We investigated five types of Li-ion batteries, each represented by their unamended global market activities:
\begin{itemize}[itemsep=0pt]
    \item Li-ion, LFP, rechargeable, prismatic
    \item Li-ion, LiMn2O4, rechargeable, prismatic
    \item Li-ion, NMC111, rechargeable, prismatic
    \item Li-ion, NMC811, rechargeable, prismatic
    \item Li-ion, NCA, rechargeable, prismatic
\end{itemize}

>>>>

In addition to the waste categories, T-reX can incorporate any number of material demand categories, the defaults being based on the EU Critical Raw Materials (CRM) list for 2023~\citep{eu2023crmstudy}, which contains 30 materials considered essential to the EU economy and also at risk of supply disruption. Especially valuable in the case of CRMs is T-reX ability to uncover and quantify these threads of background consumption, which can be inconspiciusly embededded in products or consumed in seemingly distant supply chain activities. The identity of the materials considered and their categorical groupings are easily customisable by the user, and in further materials of interest to the authors were added to the example Li-ion case study, including helium, electricity, petroleum, sand, water, and natural gas. The full list of 59 materials included in the default configuration is provided in the supplementary material in ~\autoref{sec:supplementary}.

@@

\begin{figure}[H]
    \centering
    \caption{Workflow of T-reX. Application of T-reX to one or more LCA databases creates a \texttt{brightway} project with customised methods and activities that can be used to calculate waste and material inventory footprints of the activity's supply chain.}
    \includegraphics[width=14cm]{figures/T-reX_method.pdf}\label{fig:methods-flowchart}
\end{figure}

@@

We investigated five types of Li-ion batteries, each represented by their unamended global market activities:
\begin{itemize}[itemsep=0pt]
    \item Li-ion, LFP, rechargeable, prismatic
    \item Li-ion, LiMn\(_2\)O\(_4\), rechargeable, prismatic
    \item Li-ion, NMC111, rechargeable, prismatic
    \item Li-ion, NMC811, rechargeable, prismatic
    \item Li-ion, NCA, rechargeable, prismatic
\end{itemize}

### 2024-04-02 latex_files/results.tex
update

%!!!!!!!!!!!
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{figures/T-reX\_NMC811\_WasteTotalSolid.pdf}
    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
\end{figure}

@@

%! total
\autoref{fig:waste_total} (a) shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model. The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced. The LiMn2O4 battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery. In each case, there was a slight downward trend in the waste footprints between 2020 and 2100. This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion). For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\linewidth]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
    \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only). The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
\end{figure}
%!!!!

@@

\autoref{fig:top_contribution} shows the contribution of the `top-processes' to the cobalt footprint of the LiMn2O4 battery under the baseline scenario from 2020--2100. The total footprint is seen to almost triple, from 2.2~kg/kg in 2020 to 6.2~kg/kg in 2100. This result is likely a reflection of the electrification of the transport sector that is included in the REMIND model. The fractional contributions of the top processes remain relatively steady over the coming century in this case.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{figures/top-processes.png}
    \caption{Contribution of `top-processes' to the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model. [TO BE IMPROVED!]}\label{fig:top_contribution}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{figures/cpc_contribution.png
    }
    \caption{Contribution of industrial sectors to the liquid waste footprint of the NCA battery from 2020 to 2100 under the SSP2 scenario using the PkBudg500 RCP of the REMIND model. [TO BE IMPROVED!]}\label{fig:cpc_contribution}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.8\linewidth}
        \includegraphics[width=\linewidth]{figures/comparison_methods2.png}
        \caption{}\label{fig:comparison_methods_a}
    \end{subfigure}

    \begin{subfigure}[b]{0.8\linewidth}
        \includegraphics[width=\linewidth]{figures/comparison_methods1.png}
        \caption{}\label{fig:comparison_methods_b}
    \end{subfigure}
    \caption{Comparison of LCIA method `EDIP 2003 - coal no LT' (a) with the `T-reX - Coal (black)' (b) in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.[TO BE IMPROVED!]}\label{fig:comparison_methods}
\end{figure}

>>>>

%!!!!!!!!!!!
\begin{figure}[H]
    \centering
    \includegraphics[width=19cm]{figures/T-reX_LFP_chromium.png}
    \caption{A Sankey diagram of the total solid waste footprint flows for the NMC 811 battery at the global market in the \texttt{ecoinvent} database version 3.9.1 [TO BE IMPROVED!]}\label{fig:sankey_waste}
\end{figure}

@@

%! total
\autoref{fig:waste_total} (a) shows the `total solid waste' inventory footprint for the five Li-ion batteries in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model. The NMC811 battery has the largest footprint, producing over 50~kg of waste per kilogram of battery produced. The  LiMn\(_2\)O\(_4\) battery has the smallest footprint, producing less than 4~kg of waste per kilogram of battery. In each case, there was a slight downward trend in the waste footprints between 2020 and 2100. This is most notable in the period between 2020 and 2040 and is attributable to the relatively rapid decrease in fossil-fuel use that is factored into the models over this time (since many produce relatively large amounts of waste during extraction and combustion). For the total waste generated by these batteries, there was very little difference observed between the baseline and PkBudg500 RCPs.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=19cm]{figures/T-reX-wastefootprint-totalANDcarbon.pdf}
    \caption{T-reX calculated supply chain inventories for the five Li-ion batteries in the case study, showing their `footprints' for: (a) total solid waste, and (b) carbon dioxide waste (from carbon capture and storage (CCS) only). The footprints were modelled with prospective LCA databases from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model.}\label{fig:waste_totalANDcarbon}
\end{figure}
%!!!!

@@

\autoref{fig:top_contribution} shows the contribution of the `top-processes' to the graphite footprint of the  LiMn\(_2\)O\(_4\) battery under the baseline scenario from 2020--2100. The total footprint is seen to almost triple, from 2.2~kg/kg in 2020 to 6.2~kg/kg in 2100. This result is likely a reflection of the electrification of the transport sector that is included in the REMIND model. The fractional contributions of the top processes remain relatively steady over the coming century in this case.

\begin{figure}[H]
    \centering
    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-processcontributions.pdf}
    \caption{Contribution by the `top-processes' to graphite demand inventory of the LiMn\(_2\)O\(_4\) battery from 2020 to 2100 under the SSP2 scenario using the baseline RCPs of the REMIND model.}\label{fig:top_contribution}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=9cm]{figures/T-reX-wastefootprint-sectorcontributions.pdf}
    \caption{Contribution of industrial sectors to the natural gas footprint of the LFP battery from 2020 to 2100 under the SSP2 scenario using the baseline RCP of the REMIND model.}\label{fig:cpc_contribution}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=18cm]{figures/T-reX-coalANDedip.pdf}
\caption{Comparison of the (a) `T-reX - Coal (black)' demand footprint with the (b) conventional LCIA method `EDIP 2003 - coal no LT'  in the case study from 2020 to 2100 under the SSP2 scenario using the baseline and PkBudg500 RCPs of the REMIND model for all five batteries.}\label{fig:comparison_methods}
\end{figure}
